A Two-Infall Star Formation History with Outflows and Radial Migration

### 2025-02-03 src/tex/ms.tex

\todo{Copied from last paper.} We use the multi-zone GCE model tools in \vice developed by \citet{johnson_stellar_2021}. The basic setup of our models follows theirs. The disk is divided into concentric rings of width $\delta R_{\rm gal}=100$ pc. Stellar populations migrate radially under the prescription we describe in Appendix C, but each ring is otherwise described by a conventional one-zone GCE model with instantaneous mixing (see discussion in Section \ref{sec:onezone-results}). Following \citet{johnson_stellar_2021}, we do not implement radial gas flows \citep[e.g.,][]{LaceyFall1985-RadialGasFlows,BilitewskiSchonrich2012-RadialFlows}. Stellar populations are also assigned a distance from the midplane according to their age and final radius as described in Appendix C.

\todo{Copied from last paper.} We run our models with a time-step size of $\Delta t=10$ Myr up to a maximum time of $t_{\rm final}=13.2$ Gyr. Following \citet{johnson_stellar_2021}, we set \vice to form $n=8$ stellar populations per ring per time-step, and we set a maximum star-formation radius of $R_{\rm SF} = 15.5$ kpc, such that $\dot\Sigma_\star=0$ for $R_{\rm gal}>R_{\rm SF}$. The model has a full radial extent of 20 kpc, allowing a purely migrated population to arise in the outer 4.5 kpc. We adopt continuous recycling, which accounts for the time-dependent return of mass from all previous generations of stars \citep[see Equation 2 from][]{JohnsonWeinberg2020-Starbursts}. We summarize these parameters in Table \ref{tab:multizone-parameters}.

@@

\begin{table}
    \centering
    \caption{Nucleosynthetic yield sets.}
    \label{tab:yields}
    \include{output/yields.tex}
\end{table}

@@

\begin{table*}
    \centering
    \caption{Sample selection parameters and median uncertainties from APOGEE DR17 (see Section \ref{sec:observational-sample}).}
    \label{tab:sample}
    \begin{tabular}{lll}
        \hline\hline
        \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Range or Value} & \multicolumn{1}{c}{Notes} \\
        \hline
        $\log g$            & $1.0 < \log g < 3.8$          & Select giants only \\
        $T_{\rm eff}$       & $3500 < T_{\rm eff} < 5500$ K & Reliable temperature range \\
        $S/N$               & $S/N > 80$                    & Required for accurate stellar parameters \\
        ASPCAPFLAG Bits     & $\notin$ 23                   & Remove stars flagged as bad \\
        EXTRATARG Bits      & $\notin$ 0, 1, 2, 3, or 4     & Select main red star sample only \\
        Age                 & $\sigma_{\rm Age} < 40\%$     & Age uncertainty from \citetalias{Leung2023-Ages} \\
        $R_{\rm gal}$     & $3 < R_{\rm gal} < 15$ kpc    & Eliminate bulge \& extreme outer-disk stars \\
        $|z|$               & $|z| < 2$ kpc                 & Eliminate halo stars \\
        \hline
    \end{tabular}
\end{table*}

@@

\begin{figure*}
    \centering
    \includegraphics{src/tex/figures/ofe_df_comparison.pdf}
    \caption{Stellar [O/Fe] distributions produced by multi-zone models (a--d) and as observed by APOGEE (d). Stars are binned by Galactocentric radius, represented by the color scale, and absolute midplane distance $|z|$, represented by the different rows. Each distribution is normalized so that the area under the curve is 1, and the vertical scale is consistent across each row. The median APOGEE abundance uncertainties were forward-modeled onto the model outputs. For visual clarity, each distribution is smoothed with a box-car of width 0.05 dex. Model (a) assumes the fiducial parameters and the {\tt yZ1} yield set, whereas models (b--d) assume the {\tt yZ2} yield set. Model (c) also assumes a power-law SN Ia DTD, while model (d) assumes the infalling gas has metallicity ${\rm [O/H]}_{\rm CGM}={\rm [Fe/H]}_{\rm CGM}=-0.7$. The distributions from APOGEE DR17 are plotted in column (e) for reference.}
    \label{fig:ofe-df}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics{figures/abundance_evolution.pdf}
    \caption{Stellar age--abundance relations produced by select multi-zone models. Each point represents a stellar population drawn from the Solar neighborhood near the mid-plane ($7\leq R_{\rm gal}\leq 9$ kpc, $0\leq |z| \leq 0.5$ kpc) and is color-coded by its birth radius. For visual clarity, we plot a only random mass-weighted sample of \num{10000} points in each panel. The dashed line plots the ISM abundance at $R_{\rm gal}=8$ kpc over time. The black points plot the median abundance of stars in 2 Gyr-wide age bins, and the vertical error bars represent the 16th and 84th percentiles of the age distribution. Similarly, the red points with error bars plot the median APOGEE abundances for stars with age estimates from \citet{leung_variational_2023}. Each column shows results from a different multi-zone model: {\bf (a)} our fiducial model, with $y/Z_\odot=1$, $\sigma_{\rm RM8}=2.7$ kpc, and pristine gas infall; {\bf (b)} a model with higher yields $y/Z_\odot=2$; {\bf (c)} a model with greater radial migration strength $\sigma_{\rm RM8}=5$ kpc; and {\bf (d)} a model which assumes the infalling gas has metallicity ${\rm [O/H]}_{\rm CGM}={\rm [Fe/H]}_{\rm CGM}=-0.7$. {\bf Key takeaway:} the fiducial model shows two discrepancies with the data: a major dilution event at a lookback time of $\sim9$ Gyr and non-zero abundance evolution at late times. Only model (b) with higher yields can mitigate both of these issues.}
    \label{fig:abundance-evolution}
\end{figure*}

>>>>

\todo{Copied from last paper.} We use the multi-zone GCE model tools in \vice developed by \citet{johnson_stellar_2021}. The basic setup of our models follows theirs. The disk is divided into concentric rings of width $\delta R_{\rm gal}=100$ pc. Stellar populations migrate radially under the prescription we describe in Appendix C, but each ring is otherwise described by a conventional one-zone GCE model with instantaneous mixing. Following \citet{johnson_stellar_2021}, we do not implement radial gas flows \citep[e.g.,][]{lacey_chemical_1985,bilitewski_radial_2012}. Stellar populations are also assigned a distance from the midplane according to their age and final radius as described in Appendix C.

\todo{Copied from last paper.} We run our models with a time-step size of $\Delta t=10$ Myr up to a maximum time of $t_{\rm final}=13.2$ Gyr. Following \citet{johnson_stellar_2021}, we set \vice to form $n=8$ stellar populations per ring per time-step, and we set a maximum star-formation radius of $R_{\rm SF} = 15.5$ kpc, such that $\dot\Sigma_\star=0$ for $R_{\rm gal}>R_{\rm SF}$. The model has a full radial extent of 20 kpc, allowing a purely migrated population to arise in the outer 4.5 kpc. We adopt continuous recycling, which accounts for the time-dependent return of mass from all previous generations of stars \citep[see Equation 2 from][]{johnson_impact_2020}. We summarize these parameters in Table \ref{tab:multizone-parameters}.

@@

\begin{table}
    \centering
    \caption{Nucleosynthetic yield sets.}
    \include{output/yields.tex}
    \label{tab:yields}
\end{table}

@@

\begin{table*}
    \centering
    \caption{Sample selection parameters and median uncertainties from APOGEE DR17 (see Section \ref{sec:observational-sample}).}
    \label{tab:sample}
    \begin{tabular}{lll}
        \hline\hline
        \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Range or Value} & \multicolumn{1}{c}{Notes} \\
        \hline
        $\log g$            & $1.0 < \log g < 3.8$          & Select giants only \\
        $T_{\rm eff}$       & $3500 < T_{\rm eff} < 5500$ K & Reliable temperature range \\
        $S/N$               & $S/N > 80$                    & Required for accurate stellar parameters \\
        ASPCAPFLAG Bits     & $\notin$ 23                   & Remove stars flagged as bad \\
        EXTRATARG Bits      & $\notin$ 0, 1, 2, 3, or 4     & Select main red star sample only \\
        Age                 & $\sigma_{\rm Age} < 40\%$     & Age uncertainty from \citetalias{leung_variational_2023} \\
        $R_{\rm gal}$     & $3 < R_{\rm gal} < 15$ kpc    & Eliminate bulge \& extreme outer-disk stars \\
        $|z|$               & $|z| < 2$ kpc                 & Eliminate halo stars \\
        \hline
    \end{tabular}
\end{table*}

\begin{table*}
\centering
\caption{Number of APOGEE stars in each Galactic region.}
\label{tab:apogee-regions}
\input{output/apogee_regions_table.tex}
\end{table*}

@@

\begin{figure*}
    \centering
    \includegraphics{src/tex/figures/ofe_df_comparison.pdf}
    \caption{Stellar [O/Fe] distributions produced by multi-zone models (a--d) and as observed by APOGEE (e). Stars are binned by Galactocentric radius, represented by the color scale, and absolute midplane distance $|z|$, represented by the different rows. Each distribution is normalized so that the area under the curve is 1, and the vertical scale is consistent across each row. The median APOGEE abundance uncertainties were forward-modeled onto the model outputs. For visual clarity, each distribution is smoothed with a box-car of width 0.05 dex. Model (a) assumes the fiducial parameters and the {\tt yZ1} yield set, whereas models (b--d) assume the {\tt yZ2} yield set. Model (c) also assumes a power-law SN Ia DTD, while model (d) assumes the infalling gas has metallicity ${\rm [O/H]}_{\rm CGM}={\rm [Fe/H]}_{\rm CGM}=-0.7$. The distributions from APOGEE DR17 are plotted in column (e) for reference.}
    \label{fig:ofe-df}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics{figures/abundance_evolution.pdf}
    \caption{Stellar age--abundance relations produced by select multi-zone models. Each point represents a stellar population drawn from the Solar neighborhood near the midplane ($7\leq R_{\rm gal}\leq 9$ kpc, $0\leq |z| \leq 0.5$ kpc) and is color-coded by its birth radius. For visual clarity, we plot a only random mass-weighted sample of \num{10000} points in each panel. The black curve plots the ISM abundance at $R_{\rm gal}=8$ kpc over time. The red line segments plot the median abundance for APOGEE stars in 2 Gyr-wide age bins, and the shaded regions represent the 16th--84th percentiles in each bin. Each column shows results from a different multi-zone model: {\bf (a)} our fiducial model, with $y/Z_\odot=1$, $\sigma_{\rm RM8}=2.7$ kpc, and pristine gas infall; {\bf (b)} a model with higher yields $y/Z_\odot=2$; {\bf (c)} a model with greater radial migration strength $\sigma_{\rm RM8}=5$ kpc; and {\bf (d)} a model which assumes the infalling gas has metallicity ${\rm [O/H]}_{\rm CGM}={\rm [Fe/H]}_{\rm CGM}=-0.7$. {\bf Key takeaway:} the fiducial model shows two discrepancies with the data: a major dilution event at a lookback time of $\sim9$ Gyr and non-zero abundance evolution at late times. Only model (b) with higher yields can mitigate both of these issues.}
    \label{fig:abundance-evolution}
\end{figure*}

The dilution effect discussed in \ref{sec:yields} is clearly seen in the multi-zone model results. Figure \ref{fig:abundance-evolution} shows stellar age--abundance relations for four model outputs.
